---
output:
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---
```{r}
library(tidyverse)
library(tidymodels)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(RColorBrewer)
```

```{r Read in Data}
parole <- read_csv("parole.csv")
```

```{r Recode and Rename}
parole = parole %>%
  mutate(male = as_factor(male)) %>%
  mutate(race = as_factor(race)) %>%
  mutate(state= as_factor(state)) %>%
  mutate(crime = as_factor(crime)) %>%
  mutate(multiple.offenses = as_factor(multiple.offenses)) %>%
  mutate(violator = as_factor(violator))
  
  
levels(parole$male) <- c("female", "male")
levels(parole$race) <- c("white", "other")
levels(parole$state) <- c("other", "KY", "LA", "VA")
levels(parole$crime) <- c("Other", "larceny", "drug-related", "driving-related")
levels(parole$multiple.offenses) <- c("no", "yes")
levels(parole$violator) <- c("no", "yes")


```

### Task 1: Splitting Data

```{r Split the data}
set.seed(12345)
parole_split = initial_split(parole, prop = .70, strata = violator)
train = training(parole_split)
test = testing(parole_split)

```

### Task 2: Classification Tree

```{r classification tree}
parole_recipe = recipe(violator ~. , train)

tree_model = decision_tree() %>%
  set_engine("rpart", model=TRUE) %>%
  set_mode("classification")

parole_wflow=
  workflow() %>%
  add_model(tree_model) %>%
  add_recipe(parole_recipe)

parole_fit = fit(parole_wflow, train)
```

```{r pluck the fit}
tree = parole_fit %>%
  pull_workflow_fit() %>%
  pluck("fit")

```


```{r plot the tree}
fancyRpartPlot(tree, tweak = 1.5)

```

#Task 3: Classification Example

**a 40 year old from LA with a sentence of 10 years would be classified as a "non -violator" falling under the "no" classification. Fist, I looked at the at whether the person was from KY, VA, or other , the answer was no so I moved to the right. Next, is multiple offenses =no, the answer is no because he has committed multiple offenses so I moved to the right again. Next, is the max sentence less than 13 years? The answer is yes so I moved to the left. Finally, is the time served more than 2.6 years, the answer is yes again so i moved to the left, landing at "NO" where 4% of the data falls.**

```{r Tasl 4: examine CP}
parole_fit$fit$fit$fit$cptable
```

**The optimal cp values is an xerror 1.090909** 

### Task 5: Tunning Grid

```{r create folds}
set.seed(123)
folds= vfold_cv(train, v=5)

```

```{r tunning grid}
parole_recipe2 = recipe(violator ~., train)
  
tree_model = decision_tree(cost_complexity = tune()) %>%
  set_engine("rpart", model=TRUE) %>%
  set_mode("classification")

tree_grid = grid_regular(cost_complexity(),
                         levels = 25)

parole_wflow2 =
  workflow() %>%
  add_model(tree_model) %>%
  add_recipe(parole_recipe2)

tree_res = 
  parole_wflow2 %>%
  tune_grid(
    resamples = folds,
    grid = tree_grid
  )

tree_res

```

```{r task 5: copied code to graphically examine CP}
tree_res %>%
  collect_metrics() %>%
  ggplot(aes(cost_complexity, mean)) +
  geom_line(size = 1.5, alpha = .6) +
  geom_point(size =2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2)
```

```{r task 6}
best_tree = tree_res %>%
  select_best("accuracy")

best_tree
```


**.1 yields the optimal accuracy**   


```{r task 7: finalize workflow}
final_wf =
  parole_wflow2 %>%
  finalize_workflow(best_tree)
```

```{r task 7: pluck the fit and plot the tree}
final_fit = fit(final_wf, train)

tree = final_fit %>%
  pull_workflow_fit() %>%
  pluck("fit")

#fancyRpartPlot(tree, tweak = 1.5)
```

**the accuracy of the root is is 88%**  

### Task 9: Blood Dataset

```{r Task 9 a: read in data and recode}
Blood = read_csv("Blood.csv")

Blood = Blood %>%
  mutate(DonatedMarch = as_factor(DonatedMarch))
  levels(Blood$DonatedMarch) = c("No" , "Yes")
```

```{r Task 9b: split the data}

set.seed(1234)
blood_split = initial_split(Blood, prop = .70, strata = DonatedMarch)
train2 = training(blood_split)
test2 = testing(blood_split)
```

```{r Task 9b: Classification Tree}
blood_recipe = recipe(DonatedMarch ~. , train2)

tree_model2 = decision_tree() %>%
  set_engine("rpart", model=TRUE) %>%
  set_mode("classification")

blood_wflow=
  workflow() %>%
  add_model(tree_model2) %>%
  add_recipe(blood_recipe)

blood_fit = fit(blood_wflow, train2)
```


```{r Task 9b: Create Folds}
set.seed(1234)
folds2= vfold_cv(train2, v=5)

```

```{r Task9b: Tunning}
blood_recipe2 = recipe(DonatedMarch ~., train2)
  
tree_model3 = decision_tree(cost_complexity = tune()) %>%
  set_engine("rpart", model=TRUE) %>%
  set_mode("classification")

tree_grid2 = grid_regular(cost_complexity(),
                         levels = 25)

blood_wflow2 =
  workflow() %>%
  add_model(tree_model3) %>%
  add_recipe(blood_recipe2)

tree_res2 = 
  blood_wflow2 %>%
  tune_grid(
    resamples = folds2,
    grid = tree_grid2
  )

tree_res2

```

```{r task 9b: Graphically examine CP}
tree_res2 %>%
  collect_metrics() %>%
  ggplot(aes(cost_complexity, mean)) +
  geom_line(size = 1.5, alpha = .6) +
  geom_point(size =2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2)
```


**right around .015 seems to be the best accuracy point, when examined graphically**

```{r Task 10 : finalize the workflow}
best_tree2 = tree_res2 %>%
  select_best("accuracy")

best_tree2

final2_wf =
  blood_wflow %>%
  finalize_workflow(best_tree2)
```

```{r task 10: Plot the tree}
final2_fit = fit(final2_wf, train2)

tree2 = final2_fit %>%
  pull_workflow_fit() %>%
  pluck("fit")

fancyRpartPlot(tree2, tweak = 1.5)

```



```{r Task 11: determine accuracy on training set}

treepred = predict(final2_fit, train2, type = "class")
head(treepred)

confusionMatrix(treepred$.pred_class, train2$DonatedMarch, positive = "Yes")
```

```{r determine accuracy on testing set}

treepred_test =predict(final2_fit, test2, type="class")
confusionMatrix(treepred_test$.pred_class, test2$DonatedMarch, positive = "Yes")
```

**the training set has an accuracy of .813 and the testing test an accuracy of .7857**  
