---
title: "R and RStudio Refresher"
author: "Alejandra Madrid"
date: "1/20/2021"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Task 1

```{r Loading Packages}
library(tidyverse)
library(ggplot2)
```

#Task 2

```{r}
diamonddata=diamonds
```

#Task 3
```{r}
ggplot(diamonddata, aes(carat,price)) +
geom_point()  
```
  **From this chart, we can observe that the price of a diamond increases as the size of the carat increases.**

#Task 4
```{r}
ggplot(diamonddata, aes(carat, price, color=cut)) +
  geom_point()  
```
  **From this chart, we can determine only one thing. Cut isnt indicative of price or carat. We can see the ideal cut is all over the chart, from a small carat, to a low price. We can also conclude that bigger carat does not necessarily mean better, as the bigger the carat the more likely it is that it will  fair cut.**
  
  # Task 5
```{r}
ggplot(diamonddata, aes(carat,price, color=cut)) +
  geom_point() +
  facet_wrap(~ color)  
```
  **This chart demonstrates the different prices and carat sizes there are within each color diamond. For example, that H, I, and J have bigger carat sizes and their distribution spread further. While the other call seem to be mostly carat size 3 or less.**

#Task 6 - 8 
```{r}
InventoryData <- read_csv("InventoryData.csv")
inventoryA <- InventoryData %>%
  filter(Supplier =="A")

inventoryA =mutate(inventoryA, OnHandRatio =`On Hand`/`Annual Demand`)
```
  **There are 3,695 rows.**  
**The mutate verb creates another variable/column named "OnHandRatio, that is determined by the "on Hand" divided by "Annual Demand".** 

#Task 9 
```{r}
avg_cost <- inventoryA %>%
  group_by(`Item SKU`) %>%
  summarize(SKUAvgCost = mean(`Cost per Unit ($)`))
```

#Task 10

**To me, task 9 was the hardest. It required a lot more thinking than the previous exercises. It was alot of fun, and very helpful. I think moving forward I will be alot more prepared since this exercise helped me.**